---
title: "Topic 5: Correlations"
author: "Jeroen Claes | <jeroen.claes@kuleuven.be> | <jeroen@cropland.be>"
output: 
 html_document:
  self_contained: false 
---

```{r, include=FALSE}
#library(testwhat)
tutorial::go_interactive()
```


## 1. The relationship between `Word length` and  `age of acquisition` 
- Kuperman et al. (2012) gathered age of acquisition estimates for a large collection of words via Amazon Mechanical Turk
- Here we use a sample of 1000 words
- Resarch question:
    - *Are shorter nouns acquired earlier on in life?*
- Hypothesis:
    - *Word length correlates positively with age of acquisition*
- Null hypothesis:
    - *Word length is not correlated with age of acquisition*


### 1.1 Loading and exploring the data
```{r ex="Kuperman_load_explore_associations", type="sample-code"}
# Load the readr package

# Load the course data from the course website to the object 'dataSet':
# http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Kuperman_et_al_2012.csv

# Load the dplyr package

# Print a 'glimpse' of the dataSet

# Print a summary of the AgeOfAcquisition column

# Print a summary of the WordLength column

```

```{r ex="Kuperman_load_explore_associations", type="solution"}
# Load the readr package
library(readr)
# Load the course data from the course website to the object 'dataSet':
# http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Kuperman_et_al_2012.csv
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Kuperman_et_al_2012.csv")
# Load the dplyr package
library(dplyr)
# Print a 'glimpse' of the dataSet
glimpse(dataSet)
# Print a summary of the AgeOfAcquisition column
summary(dataSet$AgeOfAcquisition)
# Print a summary of the WordLength column
summary(dataSet$WordLength)
```

```{r ex="Kuperman_load_explore_associations", type="sct"}
test_object("dataSet")
test_library_function("readr", "Make sure to call the 'readr' package!")
test_library_function("dplyr", "Make sure to call the 'dplyr' package!")
test_output_contains("glimpse(dataSet)",   incorrect_msg = "Make sure to print a 'glimpse' of the data!")
test_output_contains("summary(dataSet$AgeOfAcquisition)",   incorrect_msg = "Make sure to print a 'summary' of the AgeOfAcquisition column!")
test_output_contains("summary(dataSet$WordLength)",   incorrect_msg = "Make sure to print a 'summary' of the WordLength column!")
success_msg("Great!")
```

### 1.2  A first visual inspection of the relationship
```{r ex="Kuperman_boxplot_associations", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Kuperman_et_al_2012.csv")
```

```{r ex="Kuperman_boxplot_associations", type="sample-code"}
# The data.frame dataSet is already in your workspace

# Load the ggplot2 package

# Draw a scatterplot of  WordLength (X-axis) vs AgeOfAcquisition (y-axis)
# Add a linear (lm) trendline 

# What do you see? Is the relationship monotonic and linear? You will find the correct answer on the 'Solution' tab

```

```{r ex="Kuperman_boxplot_associations", type="solution"}
# The data.frame dataSet is already in your workspace

# Load the ggplot2 package
library(ggplot2)
# Draw a scatterplot of  WordLength (X-axis) vs AgeOfAcquisition (y-axis)
# Add a linear (lm) trendline 
ggplot(dataSet, aes(x=WordLength, y=AgeOfAcquisition)) + 
  geom_point() + 
  geom_smooth(method="lm")
# What do you see? Is the relationship monotonic and linear? 
# The relationship is monotonic: When word length goes up, there is also an increase in AgeOfAcquisition
# The relationship is not linear: When word length goes up with one unit, there is not a constant increase in the values of AgeOfAcquisition across the word length continuum
```

```{r ex="Kuperman_boxplot_associations", type="sct"}
test_library_function("ggplot2", "Make sure to call the 'ggplot2' package!")
test_ggplot(index = 1)
success_msg("Great!")
```

### 1.3  Testing the assumptions of Pearson's *r*: bivariate normal distribution and smaple size
- The plot tells us that the relationship is not linear, but it is monotonic
- This tells us already that we cannot use the Pearson's correlation without transforming the data
- For practice, let's see if the data satisfy the other assumptions 

```{r ex="Kuperman_bivariate_normal", type="pre-exercise-code"}
library(readr)
library(energy)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Kuperman_et_al_2012.csv")
```

```{r ex="Kuperman_bivariate_normal", type="sample-code"}
# The data.frame dataSet is already in your workspace

# Load the energy package

# Perform a mvnorm.etest to see if 'AgeOfAcquisition'  and WordLength have a bivariate normal distribution. Use 1000 replicates

# Check the sample size by computing the number of rows in dataSet

# What does the test tell you, do the two columns have a bivariate normal distribution? You will find the correct answer on the 'Solution' tab
```

```{r ex="Kuperman_bivariate_normal", type="solution"}
# The data.frame dataSet is already in your workspace

# Load the energy package
library(energy)

# Perform a mvnorm.etest to see if 'AgeOfAcquisition'  and WordLength have a bivariate normal distribution. Use 1000 replicates
mvnorm.etest(dataSet[, c("AgeOfAcquisition", "WordLength")], 1000)

# Check the sample size by computing the number of rows in dataSet
nrow(dataSet)

# What does the test tell you, do the two columns have a bivariate normal distribution?
# p < 0.05, so the data do not follow a bivariate normal distribution. Note that this is not that big of an issue, because the sample size is large
```

```{r ex="Kuperman_bivariate_normal", type="sct"}
test_library_function("energy", "Make sure to call the 'energy' package!")
test_output_contains('mvnorm.etest(dataSet[, c("AgeOfAcquisition", "WordLength")], 1000)', "Make sure you perform the mvnorm.etest!")
test_output_contains('nrow(dataSet)', "Don't forget to compute the sample size!")
success_msg("Excellent work!")
```

### 1.4  Testing the assumptions of Pearson's *r*: homoskedastic data

```{r ex="Kuperman_heteroske_normal", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Kuperman_et_al_2012.csv")
```

```{r ex="Kuperman_heteroske_normal", type="sample-code"}
# The data.frame dataSet is already in your workspace

# Load the car package

# Specificy a linear regression model 'mod' that regresses AgeOfAcquisition on WordLength 

# Perform a ncvTest on this regression model

# What does the test tell you, are the data heteroskedastic or homoskedastic? You will find the correct answer on the 'Solution' tab
```

```{r ex="Kuperman_heteroske_normal", type="solution"}
# The data.frame dataSet is already in your workspace

# Load the car package
library(car)
# Specificy a linear regression model 'mod' that regresses AgeOfAcquisition on WordLength 
mod <- lm(AgeOfAcquisition ~ WordLength, dataSet)
# Perform a ncvTest on this regression model
ncvTest(mod)
# What does the test tell you, are the data heteroskedastic or homoskedastic?
# p < 0.05 confirms what we saw on the plot: the relationship is not linear and the data are heteroskedastic

```

```{r ex="Kuperman_heteroske_normal", type="sct"}
test_library_function("car", "Make sure to call the 'car' package!")
test_output_contains('ncvTest(mod)', "Make sure you perform the ncvTest!")
success_msg("Great!")
```


### 1.5  Testing the assumptions of Pearson's *r*: independent data

```{r ex="Kuperman_autocorr_normal", type="pre-exercise-code"}
library(readr)
library(energy)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Kuperman_et_al_2012.csv")
```

```{r ex="Kuperman_autocorr_normal", type="sample-code"}
# The data.frame dataSet is already in your workspace

# Load the car package

# Specificy a linear regression model 'mod' that regresses AgeOfAcquisition on WordLength 

# Perform a Durbin-Watson test on this regression model

# What does the test tell you, are the data autocorrelated? You will find the correct answers on the Solution tab
```

```{r ex="Kuperman_autocorr_normal", type="solution"}
# The data.frame dataSet is already in your workspace

# Load the car package
library(car)
# Specificy a linear regression model 'mod' that regresses AgeOfAcquisition on WordLength 
mod <- lm(AgeOfAcquisition ~ WordLength, dataSet)
# Perform a ncvTest on this regression model
durbinWatsonTest(mod)
# What does the test tell you, are the data autocorrelated?
# p > 0.05 tells us that autocorrelation is not an issue
```

```{r ex="Kuperman_autocorr_normal", type="sct"}
test_library_function("car", "Make sure to call the 'car' package!")
test_output_contains('durbinWatsonTest(mod)', "Make sure you perform the durbinWatsonTest!")
success_msg("Great!")
```

### 1.6  Performing the correlations test
- All of our previous exercises point in the same direction: the data are not suited for the Pearson's correlation test:
    - The relationship is not linear
    - The relationship is heteroskedastic
    - The data do not follow a bivariate normal distribution (less important, as the sample size is large)
- Spearman's *Rho* and Kendall's *tau* are our only two options

```{r ex="Kuperman_test_normal", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Kuperman_et_al_2012.csv")
```

```{r ex="Kuperman_test_normal", type="sample-code"}
# The data.frame dataSet is already in your workspace

# Perform a cor.test with the 'spearman' method. Recall that our hypothesis is directional: it predicts a positive correlation

# Perform a cor.test with the 'kendall' method. Recall that our hypothesis is directional: it predicts a positive correlation

# What do the two tests tell you: 
# - Is the relationship between the two variables strong? 
# - Is the relationship positive or negative? 
# - Is it significant? 
# Which test yields the most extreme estimate?
# You will find the correct answers on the Solution tab
```

```{r ex="Kuperman_test_normal", type="solution"}
# The data.frame dataSet is already in your workspace

# Perform a cor.test with the 'spearman' method. Recall that our hypothesis is directional: it predicts a positive correlation
cor.test(dataSet$AgeOfAcquisition, dataSet$WordLength, alternative="greater", method="spearman")

# Perform a cor.test with the 'kendall' method. Recall that our hypothesis is directional: it predicts a positive correlation
cor.test(dataSet$AgeOfAcquisition, dataSet$WordLength, alternative="greater", method="kendall")

# What does the two tests tell you: 
# - Is the relationship between the two variables strong? 
# The relationship is weak to moderate (rho and tau around 0.3)
# - Is the relationship positive or negative? 
# The relationship is positive, the coefficients are greater than zero
# - Is it significant? 
# p < 0.05 so we can consider the relationship to be weak, but significant
# Which test yields the most extreme estimate?
# As is often the case, Spearman's rho is more extreme than Kendall's tau
# You will find the correct answers on the Solution tab
```

```{r ex="Kuperman_test_normal", type="sct"}
test_output_contains('cor.test(dataSet$AgeOfAcquisition, dataSet$WordLength, alternative="greater", method="kendall")', "Make sure you perform the cor.test with the 'kendall' method!")
test_output_contains('cor.test(dataSet$AgeOfAcquisition, dataSet$WordLength, alternative="greater", method="spearman")', "Make sure you perform the cor.test with the 'spearman' method!")
success_msg("Good job!")
```

## 2. The relationship between Degree of `synthesis` and  `isolation` across languages 
- Greenberg (1960) calculated different indices based on a large collection of texts for different languages (the data were donated by Freek Van de Velde, KU Leuven):
    - `Synthesis_Index` indicator of the degree of syntheticity of a language. It expresses the ratio of morphemes to words (i.e., how many morphemes occur in a typical word). Ranges from 0 to +Inf, but values above 3 do not occur
    - `Isolation_Index` indicator of the number of words that are necessary to convey a relationship between words at the level of the sentence. Ranges from 0 (no words) to 1 (one word)
    - Since the two indices are on a different scale, they were transformed to z-scores
- Our hypothesis is the following:
    - *Languages high on the Synthesis index will be low on the Isolation index and vice versa* 
- Our null hypothesis states:
    - *There is no relationship between syntheticity and isolation in language*

### 2.1 Loading and exploring the data
```{r ex="Greenberg_load_explore_associations", type="sample-code"}
# Load the readr package

# Load the course data from the course website to the object 'dataSet':
# http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Greenberg_1960.csv

# Load the dplyr package

# Print a 'glimpse' of the dataSet

# Print a summary of the Synthesis_Index column

# Print a summary of the Isolation_Index column

```

```{r ex="Greenberg_load_explore_associations", type="solution"}
# Load the readr package
library(readr)
# Load the course data from the course website to the object 'dataSet':
# http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Greenberg_1960.csv
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Greenberg_1960.csv")
# Load the dplyr package
library(dplyr)
# Print a 'glimpse' of the dataSet
glimpse(dataSet)
# Print a summary of the Synthesis_Index column
summary(dataSet$Synthesis_Index)
# Print a summary of the Isolation_Index column
summary(dataSet$Isolation_Index)
```

```{r ex="Greenberg_load_explore_associations", type="sct"}
test_object("dataSet")
test_library_function("readr", "Make sure to call the 'readr' package!")
test_library_function("dplyr", "Make sure to call the 'dplyr' package!")
test_output_contains("glimpse(dataSet)",   incorrect_msg = "Make sure to print a 'glimpse' of the data!")
test_output_contains("summary(dataSet$Synthesis_Index)",   incorrect_msg = "Make sure to print a 'summary' of the Synthesis_Index column!")
test_output_contains("summary(dataSet$Isolation_Index)",   incorrect_msg = "Make sure to print a 'summary' of the Isolation_Index column!")
success_msg("Great!")
```

### 2.2  A first visual inspection of the relationship
```{r ex="Greenberg_boxplot_associations", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Greenberg_1960.csv")
```

```{r ex="Greenberg_boxplot_associations", type="sample-code"}
# The data.frame dataSet is already in your workspace

# Load the ggplot2 package

# Draw a scatterplot of  Isolation_Index (X-axis) vs Synthesis_Index (y-axis)
# Add a linear (lm) trendline 

# What do you see? Is the relationship monotonic and linear? You will find the correct answer on the 'Solution' tab

```

```{r ex="Greenberg_boxplot_associations", type="solution"}
# The data.frame dataSet is already in your workspace

# Load the ggplot2 package
library(ggplot2)
# Draw a scatterplot of  Isolation_Index (X-axis) vs Synthesis_Index (y-axis)
# Add a linear (lm) trendline 
ggplot(dataSet, aes(x=Isolation_Index, y=Synthesis_Index)) + 
  geom_point() + 
  geom_smooth(method="lm")
# What do you see? Is the relationship monotonic and linear? 
# The relationship is monotonic: When Synthesis_Index goes up, there is a decrease in Isolation_Index
# The relationship is linear: When Synthesis_Index goes up with one unit, there is a constant decrease in the values of Isolation_Index across the Isolation_Index continuum
```

```{r ex="Greenberg_boxplot_associations", type="sct"}
test_library_function("ggplot2", "Make sure to call the 'ggplot2' package!")
test_ggplot(index = 1)
success_msg("Fantastic work!")
```

### 2.3  Testing the assumptions of Pearson's *r*: bivariate normal distribution and sample size
- The plot tells us that the relationship is linear and monotonic
- This tells us that Pearson's *r* could be a valid correlation statistic for our data
- Let's see if the data also satisfy its other assumptions 

```{r ex="Greenberg_bivariate_normal", type="pre-exercise-code"}
library(readr)
library(energy)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Greenberg_1960.csv")
```

```{r ex="Greenberg_bivariate_normal", type="sample-code"}
# The data.frame dataSet is already in your workspace

# Load the energy package

# Perform a mvnorm.etest to see if 'Synthesis_Index'  and Isolation_Index have a bivariate normal distribution. Use 1000 replicates

# Compute the sample size by calculating the number of rows in dataSet

# What does the test tell you, do the two columns have a bivariate normal distribution? You will find the correct answer on the 'Solution' tab
```

```{r ex="Greenberg_bivariate_normal", type="solution"}
# The data.frame dataSet is already in your workspace

# Load the energy package
library(energy)
# Perform a mvnorm.etest to see if 'Synthesis_Index'  and Isolation_Index have a bivariate normal distribution. Use 1000 replicates
mvnorm.etest(dataSet[, c("Synthesis_Index", "Isolation_Index")], 1000)

# Compute the sample size by calculating the number of rows in dataSet
nrow(dataSet)

# What does the test tell you, do the two columns have a bivariate normal distribution?
# p > 0.05, so the data follow a bivariate normal distribution. Note that this is important here, as the sample size is so small
```

```{r ex="Greenberg_bivariate_normal", type="sct"}
test_library_function("energy", "Make sure to call the 'energy' package!")
test_output_contains("nrow(dataSet)", "Don't forget to compute the sample size!")
test_output_contains('mvnorm.etest(dataSet[, c("Synthesis_Index", "Isolation_Index")], 1000)', "Make sure you perform the mvnorm.etest!")
success_msg("Good job!")
```

### 2.4  Testing the assumptions of Pearson's *r*: homoskedastic data

```{r ex="Greenberg_heteroske_normal", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Greenberg_1960.csv")
```

```{r ex="Greenberg_heteroske_normal", type="sample-code"}
# The data.frame dataSet is already in your workspace

# Load the car package

# Specificy a linear regression model 'mod' that regresses Synthesis_Index on Isolation_Index 

# Perform a ncvTest on this regression model

# What does the test tell you, are the data heteroskedastic or homoskedastic? You will find the correct answer on the 'Solution' tab
```

```{r ex="Greenberg_heteroske_normal", type="solution"}
# The data.frame dataSet is already in your workspace

# Load the car package
library(car)
# Specificy a linear regression model 'mod' that regresses Synthesis_Index on Isolation_Index 
mod <- lm(Synthesis_Index ~ Isolation_Index, dataSet)
# Perform a ncvTest on this regression model
ncvTest(mod)
# What does the test tell you, are the data heteroskedastic or homoskedastic?
# p > 0.05 confirms what we saw on the plot: the relationship is linear and the data are homoskedastic

```

```{r ex="Greenberg_heteroske_normal", type="sct"}
test_library_function("car", "Make sure to call the 'car' package!")
test_output_contains('ncvTest(mod)', "Make sure you perform the ncvTest!")
success_msg("Great!")
```


### 2.5  Testing the assumptions of Pearson's *r*: independent data

```{r ex="Greenberg_autocorr_normal", type="pre-exercise-code"}
library(readr)
library(energy)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Greenberg_1960.csv")
```

```{r ex="Greenberg_autocorr_normal", type="sample-code"}
# The data.frame dataSet is already in your workspace

# Load the car package

# Specificy a linear regression model 'mod' that regresses Synthesis_Index on Isolation_Index 

# Perform a Durbin-Watson test on this regression model

# What does the test tell you, are the data autocorrelated? You will find the correct answer on the Solution tab
```

```{r ex="Greenberg_autocorr_normal", type="solution"}
# The data.frame dataSet is already in your workspace

# Load the car package
library(car)
# Specificy a linear regression model 'mod' that regresses Synthesis_Index on Isolation_Index 
mod <- lm(Synthesis_Index ~ Isolation_Index, dataSet)
# Perform a ncvTest on this regression model
durbinWatsonTest(mod)
# What does the test tell you, are the data autocorrelated?
# p > 0.05 tells us that autocorrelation is not an issue
```

```{r ex="Greenberg_autocorr_normal", type="sct"}
test_library_function("car", "Make sure to call the 'car' package!")
test_output_contains('durbinWatsonTest(mod)', "Make sure you perform the durbinWatsonTest!")
success_msg("Great!")
```

### 2.6  Performing the correlations test
- All of our previous exercises point in the same direction: the data are perfectly suited for the Pearson's correlation test:
    - The relationship is linear and monotonic
    - The relationship is homoskedastic
    - The data follow a bivariate normal distribution (this is important, because the sample size is small)
- We can use Pearson's *r*

```{r ex="Greenberg_test_normal", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class5_Greenberg_1960.csv")
```

```{r ex="Greenberg_test_normal", type="sample-code"}
# The data.frame dataSet is already in your workspace

# Perform a cor.test with the 'pearson' method. Recall that our hypothesis is directional: it predicts a negative correlation

# Calculate the r-squared value of the relationship. How much of the variance of  Isolation_Index is explained by Synthesis_Index?

# What do the tests tell you: 
# - Is the relationship between the two variables strong? 
# - Is the relationship positive or negative? 
# - Is it significant? 
# - How much of the variance of  Isolation_Index is explained by Synthesis_Index?
# You will find the correct answers on the Solution tab
```

```{r ex="Greenberg_test_normal", type="solution"}
# The data.frame dataSet is already in your workspace

# Perform a cor.test with the 'pearson' method. Recall that our hypothesis is directional: it predicts a negative correlation
cor.test(dataSet$Synthesis_Index, dataSet$Isolation_Index, alternative="less", method="pearson")

# Calculate the r-squared value of the relationship. 
cor(dataSet$Synthesis_Index, dataSet$Isolation_Index)^2

# What do the tests tell you: 
# - Is the relationship between the two variables strong? 
# The relationship is very strong
# - Is the relationship positive or negative? 
# The relationship is negative, the coefficient is smaller than zero
# - Is it significant? 
# p < 0.05 so we can consider the relationship to be strong AND significant
# - How much of the variance of  Isolation_Index is explained by Synthesis_Index?
# The Synthesis_Index explains about 71% of the variance of Isolation_Index
```

```{r ex="Greenberg_test_normal", type="sct"}
test_output_contains('cor.test(dataSet$Synthesis_Index, dataSet$Isolation_Index, alternative="less", method="pearson")', "Make sure you perform the cor.test with the 'pearson' method!")
test_output_contains('cor(dataSet$Synthesis_Index, dataSet$Isolation_Index)^2', "Don't forget to calculate the r-squared value!")
success_msg("Good work!")
```


## References
- Greenberg, J. (1960). A quantitative approach to the morphological typology of language. *International Journal of American Linguistics* 26(3). 178-194.
- Kuperman, V., Stadthagen-Gonzalez,  H., Brysbaert, M. (2012). Age-of-acquisition ratings for 30,000 English words.*Behavior Research Methods* 44(4).978–990.

## Acknowledgements
A warm *thank you!* goes out to **Freek Van de Velde** (KU Leuven), who generously donated the Greenberg (1960) data for the second exercise