---
title: 'Topic 2: Data frames, measures of central tendency, measures of dispersion,
  and distributions'
author: "Jeroen Claes | <jeroen.claes@kuleuven.be> | <jeroen@cropland.be>"
---
## 1. Working with data.frames
### Loading data from CSV
```{r, include=FALSE}
tutorial::go_interactive()
```

```{r ex="create_a", type="sample-code"}
# Load the readr package

# Load the course data from the course website to the object 'dataSet':
# https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv

# Load dplyr

# Print a 'glimpse' of the dataset to inspect its structure


```

```{r ex="create_a", type="solution"}
# Load the readr package
library(readr)
# Load the course data from the course website to the object 'dataSet':
# https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")

# Load dplyr
library(dplyr)

# Print a 'glimpse' of the dataset to inspect its structure
glimpse(dataSet)
```

```{r ex="create_a", type="sct"}
test_object("dataSet")
test_library_function("dplyr")
test_library_function("readr")
test_output_contains("glimpse(dataSet)", incorrect_msg = "Make sure to print the last six rows!")
success_msg("Great!")
```

### Exploring data.frames (1/2)

```{r ex="explore", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="explore", type="sample-code"}
# The data.frame dataSet is already in your workspace. Call ls() to see it

# Explore its dimensions: how many rows, how many columns?

# Print the first 15 rows of the data.frame


```

```{r ex="explore", type="solution"}
# The data.frame dataSet is already in your workspace. Call ls() to see it
ls()
# Explore its dimensions: how many rows, how many columns?
dim(dataSet)
# Print the first 15 rows of the data.frame
head(dataSet, 15)

```

```{r ex="explore", type="sct"}
test_output_contains("dim(dataSet)", incorrect_msg = "Make sure to print the dimensions of the data.frame.")
test_output_contains("head(dataSet, 15)", incorrect_msg = "Make sure to print the first 15 lines.")
success_msg("Great!")
```

### Exploring data.frames (2/2)

```{r ex="explore2", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="explore2", type="sample-code"}
# The data.frame dataSet is already in your workspace. 

# Explore its dimensions: how many rows does it have?

# Explore its dimensions: how many columns does it have?

# Print a summary of the data.frame
```

```{r ex="explore2", type="solution"}
# The data.frame dataSet is already in your workspace. Call ls() to see it
ls()
# Explore its dimensions: how many rows does it have?
nrow(dataSet)
# Explore its dimensions: how many columns does it have?
ncol(dataSet)
# Print a summary of the data.frame
summary(dataSet)
```

```{r ex="explore2", type="sct"}
test_output_contains("nrow(dataSet)", incorrect_msg = "Make sure to print the number of rows")
test_output_contains("ncol(dataSet)", incorrect_msg = "Make sure to print the number of columns")
test_output_contains("summary(dataSet)", incorrect_msg = "Make sure to print a summary")
success_msg("Great!")
```

### Subsetting data.frames (1/2)
```{r ex="subset1", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="subset1", type="sample-code"}
# The data.frame dataSet is already in your workspace. 

# Print the last couple of values of its 'Freq' column

# Create a new data.frame, called "lowFrequency,  which holds all rows of dataSet, where Freq is smaller than or equal to 30

# Print the first, the second, and the 50th row of this data.frame


```

```{r ex="subset1", type="solution"}
# The data.frame dataSet is already in your workspace. 

# Print the last couple of values of its 'Freq' column
tail(dataSet$Freq)
# Create a new data.frame, called "lowFrequency,  which holds all rows and columns of dataSet, where Freq is smaller than or equal to 30
lowFrequency <- dataSet[dataSet$Freq <= 30,]

# Print the first, the second, and the 50th row of this data.frame
lowFrequency[c(1, 2, 50),]
```

```{r ex="subset1", type="sct"}

test_output_contains("tail(dataSet$Freq)", incorrect_msg = "Make sure to print the last couple of values of Freq")
test_object("lowFrequency")
test_output_contains("lowFrequency[c(1, 2, 50),]", incorrect_msg = "Make sure to print the first, the second, and the 50th row of lowFrequency")
success_msg("Great!")
```

### Subsetting data.frames (2/2)
```{r ex="subset2", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="subset2", type="sample-code"}
# The data.frame dataSet is already in your workspace. 

# Extract all values of Word with Mean_RT smaller than 600. Print only the first couple of words

# Extract Length and Freq of the following words: "moneys"  "stinger" "spare"   "bandage" "acorn"  

# Add a new variable to the data.frame, called FreqByTwo which stores the Freq of the word, divided by two

# Print a summary of the modified data.Frame
```

```{r ex="subset2", type="solution"}
# The data.frame dataSet is already in your workspace. 

# Extract all Words with a reaction time smaller than 600. Print only the first couple of words
head(dataSet[dataSet$Mean_RT < 600,]$Word)

# Extract the Length and Freq of the following words: "moneys"  "stinger" "spare"   "bandage" "acorn"  
dataSet[dataSet$Word %in% c("moneys", "stinger", "spare", "bandage","acorn"), c("Length", "Freq")]

# Add a new variable to the data.frame, called FreqByTwo which stores the Freq of the word, divided by two
dataSet$FreqByTwo <- dataSet$Freq/2

# Print a summary of the modified data.Frame
summary(dataSet)
```

```{r ex="subset2", type="sct"}
test_output_contains("head(dataSet[dataSet$Mean_RT < 600,]$Word)", incorrect_msg = "Make sure to print the first couple of values of Freq < 600")
test_output_contains('dataSet[dataSet$Word %in% c("moneys", "stinger", "spare", "bandage","acorn"), c("Length", "Freq")]', incorrect_msg = "Make sure to extract the Length and the Freq of the words in the list")
success_msg("Great job!")
```

## 2. Measures of central tendency
- Calculate these measures for the variables `Freq` and `Mean_RT`
- **ANALYZE**:
    - What differences do you spot between `Freq` and `Mean_RT` on the one hand and `Length` on the other?

### 2.1 The `Freq` variable

```{r ex="Freq", type="sample-code"}
# Load the data again from
# https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv
# Remember that you will need the package readr for that

# Calculate the mean of 'Freq'

# Calculate the median of 'Freq'

# Calculate the 0.25, 0.5, and 0.75 quantiles of 'Freq'

# Calculate the mode of 'Freq'

# Calculate the minimum of 'Freq'

# Calculate the maximum of 'Freq'
```


```{r ex="Freq", type="solution"}
# Load the data again from
# https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv
# Remember that you will need the package readr for that
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
# Calculate the mean of 'Freq'
mean(dataSet$Freq, na.rm=TRUE)
# Calculate the median of 'Freq'
median(dataSet$Freq, na.rm=TRUE)
# Calculate the 0.25, 0.5, and 0.75 quantiles of 'Freq'
quantile(dataSet$Freq, c(0.25, 0.5, 0.75), na.rm = TRUE)
# Calculate the mode of 'Freq'
sort(table(dataSet$Freq), decreasing = TRUE)[1]
# Calculate the minimum of 'Freq'
min(dataSet$Freq, na.rm=TRUE)
# Calculate the maximum of 'Freq'
max(dataSet$Freq, na.rm=TRUE)
```

```{r ex="Freq", type="sct"}
test_output_contains("mean(dataSet$Freq, na.rm=TRUE)", incorrect_msg = "Make sure to add 'na.rm=TRUE' to your 'mean' code!")
test_output_contains('median(dataSet$Freq, na.rm=TRUE)', incorrect_msg = "Make sure to add 'na.rm=TRUE' to your 'median' code!")
test_output_contains("quantile(dataSet$Freq, c(0.25, 0.5, 0.75), na.rm = TRUE)", incorrect_msg = "Make sure to add 'na.rm=TRUE' to your 'quantile' code!")
test_output_contains("sort(table(dataSet$Freq), decreasing = TRUE)[1]", incorrect_msg = "Make sure you use the one-line alternative to calculate the mode!")
test_output_contains("min(dataSet$Freq, na.rm=TRUE)", incorrect_msg = "Make sure you add 'na.rm=TRUE' to your 'min' code!")
test_output_contains("max(dataSet$Freq, na.rm=TRUE)", incorrect_msg = "Make sure you add 'na.rm=TRUE' to your 'max' code!")
test_library_function("readr")
success_msg("Great job!")
```

### 2.1 The `Mean_RT` variable
```{r ex="Mean_RT", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")

```

```{r ex="Mean_RT", type="sample-code"}
# The object dataSet is already in your workspace

# Calculate the mean of 'Mean_RT'

# Calculate the median of 'Mean_RT'

# Calculate the 0.25, 0.5, and 0.75 quantiles of 'Mean_RT'

# Calculate the mode of 'Mean_RT'

# Calculate the minimum of 'Mean_RT'

# Calculate the maximum of 'Mean_RT'
```


```{r ex="Mean_RT", type="solution"}
# Calculate the mean of 'Mean_RT'
mean(dataSet$Mean_RT, na.rm=TRUE)
# Calculate the median of 'Mean_RT'
median(dataSet$Mean_RT, na.rm=TRUE)
# Calculate the 0.25, 0.5, and 0.75 quantiles of 'Mean_RT'
quantile(dataSet$Mean_RT, c(0.25, 0.5, 0.75), na.rm = TRUE)
# Calculate the mode of 'Mean_RT'
sort(table(dataSet$Mean_RT), decreasing = TRUE)[1]
# Calculate the minimum of 'Mean_RT'
min(dataSet$Mean_RT, na.rm=TRUE)
# Calculate the maximum of 'Mean_RT'
max(dataSet$Mean_RT, na.rm=TRUE)
```

```{r ex="Mean_RT", type="sct"}
test_output_contains("mean(dataSet$Mean_RT, na.rm=TRUE)", incorrect_msg = "Make sure to add 'na.rm=TRUE' to your 'mean' code!")
test_output_contains('median(dataSet$Mean_RT, na.rm=TRUE)', incorrect_msg = "Make sure to add 'na.rm=TRUE' to your 'median' code!")
test_output_contains("quantile(dataSet$Mean_RT, c(0.25, 0.5, 0.75), na.rm = TRUE)", incorrect_msg = "Make sure to add 'na.rm=TRUE' to your 'quantile' code!")
test_output_contains("sort(table(dataSet$Mean_RT), decreasing = TRUE)[1]", incorrect_msg = "Make sure you use the one-line alternative to calculate the mode!")
test_output_contains("min(dataSet$Mean_RT, na.rm=TRUE)", incorrect_msg = "Make sure you add 'na.rm=TRUE' to your 'min' code!")
test_output_contains("max(dataSet$Mean_RT, na.rm=TRUE)", incorrect_msg = "Make sure you add 'na.rm=TRUE' to your 'max' code!")
success_msg("Great job!")
```

## 3. The normal distribution
- Plot and test the `Freq` and `Mean_RT` variables. 
- **Analyze**:
    - Are these variables normally distributed?
    
### 3.1 The `Freq` variable    
####  Mean and median
```{r ex="Freq_median_mean", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Freq_median_mean", type="sample-code"}
# The object dataSet is already in your workspace
# Compute the mean of Freq. 

# Compute the median of Freq. 

# Are the mean and the median close together numerically?  You will find the correct answer on the tab Solution
```

```{r ex="Freq_median_mean", type="solution"}
## The object dataSet is already in your workspace
# Compute the mean of Freq. 
mean(dataSet$Freq, na.rm=TRUE)
# Compute the median of Freq. 
median(dataSet$Freq, na.rm=TRUE)
# Are the mean and the median close together numerically? 
# No, the mean is more than ten times higher than the median

```

```{r ex="Freq_median_mean", type="sct"}
test_output_contains("mean(dataSet$Freq, na.rm=TRUE)
", incorrect_msg = "Make sure to add 'na.rm=TRUE' to your 'mean' code!")
test_output_contains("median(dataSet$Freq, na.rm=TRUE)", incorrect_msg = "Make sure to add 'na.rm=TRUE' to your 'median' code!")
```

####  Density plot 
```{r ex="Freq_density", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Freq_density", type="sample-code"}
# The object dataSet is already in your workspace
# Load the package ggplot2

# Draw up a density plot of the Freq variable

# Does the plot have a bell shape?  You will find the correct answer on the tab Solution
```

```{r ex="Freq_density", type="solution"}
# The object dataSet is already in your workspace
# Load the package ggplot2
library(ggplot2)
# Draw up a density plot of the Freq variable and add a red line at the mean
ggplot(dataSet, aes(x=Freq))  + geom_line(stat="density") + geom_vline(aes(xintercept=mean(dataSet$Freq)), color="red")
# Does the plot have a bell shape? 
# No, not at all
```

```{r ex="Freq_density", type="sct"}
test_library_function("ggplot2")
test_ggplot(index=1, all_fail_msg = "Computer says no. This plot is completely wrong", data_fail_msg = "Make sure you include the data in the ggplot call", aes_fail_msg = "Make sure you define the correct aesthetics (aes)", geom_fail_msg = "Make sure you define the correct plotting function (geom_)")
success_msg("Great!")

```

####  QQ-plot 
```{r ex="Freq_qq", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Freq_qq", type="sample-code"}
# The object dataSet is already in your workspace
# Draw up a QQ-plot of the Freq variable

# Add a QQ-line for the Freq variable

# Does the plot support that the variable is normally distributed? You may want to enlarge the plot before answering this question.  You will find the correct answer on the tab Solution
```

```{r ex="Freq_qq", type="solution"}
# The object dataSet is already in your workspace
# Draw up a QQ-plot of the Freq variable
qqnorm(dataSet$Freq)
# Add a QQ-line for the Freq variable
qqline(dataSet$Freq)
# Does the plot support that the variable is normally distributed? You may want to enlarge the plot before answering this question. 
# No, the points do not follow the qqline at all
```

```{r ex="Freq_qq", type="sct"}
test_function("qqnorm", args = "y")
test_function("qqline", args ="y")
success_msg("Great!")
```


####  Shapiro-Wilk test
```{r ex="Freq_shap", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Freq_shap", type="sample-code"}
# The object dataSet is already in your workspace
# Perform a Shapiro-Wilk test for your variable

# Does the test support that the variable is normally distributed? You will find the correct answer on the tab Solution
```

```{r ex="Freq_shap", type="solution"}
# The object dataSet is already in your workspace
# Perform a Shapiro-Wilk test for your variable
shapiro.test(dataSet$Freq)
# Does the test support that the variable is normally distributed? You will find the correct answer on the tab Solution
#  No. p < 0.05, so we can assume that the data are not normally distributed"
```

```{r ex="Freq_shap", type="sct"}
test_output_contains("shapiro.test(dataSet$Freq)", incorrect_msg = "Make sure you perform the test!")
```

### 3.2 The `Mean_RT` variable    
####  Mean and median
```{r ex="Mean_RT_median_mean", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Mean_RT_median_mean", type="sample-code"}
# The object dataSet is already in your workspace
# Compute the mean of Mean_RT. 

# Compute the median of Mean_RT. 

# Are the mean and the median close together numerically?  You will find the correct answer on the tab Solution
```

```{r ex="Mean_RT_median_mean", type="solution"}
## The object dataSet is already in your workspace
# Compute the mean of Mean_RT. 
mean(dataSet$Mean_RT, na.rm=TRUE)
# Compute the median of Mean_RT. 
median(dataSet$Mean_RT, na.rm=TRUE)
# Are the mean and the median close together numerically? 
# Yes, the difference between the mean and the median is not really large. 

```

```{r ex="Mean_RT_median_mean", type="sct"}
test_output_contains("mean(dataSet$Mean_RT, na.rm=TRUE)
", incorrect_msg = "Make sure to add 'na.rm=TRUE' to your 'mean' code!")
test_output_contains("median(dataSet$Mean_RT, na.rm=TRUE)", incorrect_msg = "Make sure to add 'na.rm=TRUE' to your 'median' code!")
```

####  Density plot 
```{r ex="Mean_RT_density", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Mean_RT_density", type="sample-code"}
# The object dataSet is already in your workspace
# Load the package ggplot2

# Draw up a density plot of the Mean_RT variable

# Does the plot have a bell shape?  You will find the correct answer on the tab Solution
```

```{r ex="Mean_RT_density", type="solution"}
# The object dataSet is already in your workspace
# Load the package ggplot2
library(ggplot2)
# Draw up a density plot of the Mean_RT variable and add a red line at the mean
ggplot(dataSet, aes(x=Mean_RT))  + geom_line(stat="density") + geom_vline(aes(xintercept=mean(dataSet$Mean_RT)), color="red")
# Does the plot have a bell shape? 
# No, not at all
```

```{r ex="Mean_RT_density", type="sct"}
test_library_function("ggplot2")
test_ggplot(index = 1)
success_msg("Great!")

```

####  QQ-plot 
```{r ex="Mean_RT_qq", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Mean_RT_qq", type="sample-code"}
# The object dataSet is already in your workspace
# Draw up a QQ-plot of the Mean_RT variable

# Add a QQ-line for the Mean_RT variable

# Does the plot support that the variable is normally distributed? You may want to enlarge the plot before answering this question.  You will find the correct answer on the tab Solution
```

```{r ex="Mean_RT_qq", type="solution"}
# The object dataSet is already in your workspace
# Draw up a QQ-plot of the Mean_RT variable
qqnorm(dataSet$Mean_RT)
# Add a QQ-line for the Mean_RT variable
qqline(dataSet$Mean_RT)
# Does the plot support that the variable is normally distributed? You may want to enlarge the plot before answering this question. 
# No, the points do not follow the qqline at all
```

```{r ex="Mean_RT_qq", type="sct"}
test_function("qqnorm", args = c("y"), incorrect_msg = "Make sure you draw the qqplot!")
test_function("qqline", args= c("y"), incorrect_msg = "Don't forget to add the qqline!")
success_msg("Great!")
```

####  Shapiro-Wilk test
```{r ex="Mean_RT_shap", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Mean_RT_shap", type="sample-code"}
# The object dataSet is already in your workspace
# Perform a Shapiro-Wilk test for your variable

# Does the test support that the variable is normally distributed? You will find the correct answer on the tab Solution
```

```{r ex="Mean_RT_shap", type="solution"}
# The object dataSet is already in your workspace
# Perform a Shapiro-Wilk test for your variable
shapiro.test(dataSet$Mean_RT)
# Does the test support that the variable is normally distributed? You will find the correct answer on the tab Solution
#  No. p < 0.05, so we can assume that the data are not normally distributed"
```

```{r ex="Mean_RT_shap", type="sct"}
test_output_contains("shapiro.test(dataSet$Mean_RT)", incorrect_msg = "Make sure you perform the test!")
success_msg("Great job!")
```

## 4. Measures of dispersion
- Calculate the Range, Variance, standard deviation, interquartile range, and  median absolute deviation of the `Freq` and `Mean_RT` variables. 
- Analyze:
    - Which one of these two is the most dispersed? I.e., which one has the least 'normal' distribution?

### 4.1 The `Freq` variable

#### Range, Variance, standard deviation, interquartile range, and  median absolute deviation 
```{r ex="Freq_dispersion", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Freq_dispersion", type="sample-code"}
# The object dataSet is already in your workspace
# Range of Freq

# Variance of Freq

# Standard deviation of Freq

# Interquartile range of Freq

# Median absolute deviation of Freq
```

```{r ex="Freq_dispersion", type="solution"}
# The object dataSet is already in your workspace
# Range of Freq
max(dataSet$Freq, na.rm=TRUE) - min(dataSet$Freq)
# Variance of Freq
var(dataSet$Freq, na.rm=TRUE)
# Standard deviation of Freq
sd(dataSet$Freq, na.rm=TRUE)
# Interquartile range of Freq
IQR(dataSet$Freq, na.rm=TRUE)
# Median absolute deviation of Freq
mad(dataSet$Freq, na.rm=TRUE)
```

```{r ex="Freq_dispersion", type="sct"}
test_output_contains("max(dataSet$Freq, na.rm=TRUE) - min(dataSet$Freq)", incorrect_msg = "Make sure you add na.rm to your max and min statements!")
test_output_contains('var(dataSet$Freq, na.rm=TRUE)', incorrect_msg = "Make sure you add 'na.rm=TRUE' to your code to calculate the variance")
test_output_contains('sd(dataSet$Freq, na.rm=TRUE)', incorrect_msg = "Make sure you add 'na.rm=TRUE' to your code to calculate the standard deviation")
test_output_contains('IQR(dataSet$Freq, na.rm=TRUE)', incorrect_msg = "Make sure you add 'na.rm=TRUE' to your code to calculate the Interquartile range")
test_output_contains('var(dataSet$Freq, na.rm=TRUE)', incorrect_msg = "Make sure you add 'na.rm=TRUE' to your code to calculate the mean absolute deviation")
success_msg("Great job!")
```

#### Box-and-whiskers plot
```{r ex="Freq_boxplot", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Freq_boxplot", type="sample-code"}
# The object dataSet is already in your workspace
# Load ggplot2

# Draw a boxplot of Freq

```

```{r ex="Freq_boxplot", type="solution"}
# The object dataSet is already in your workspace
# Load ggplot2
library(ggplot2)
# Draw the boxplot
ggplot(dataSet, aes(x=1, y=Freq)) + geom_boxplot()
```

```{r ex="Freq_boxplot", type="sct"}
test_library_function("ggplot2")
test_ggplot(index=1, all_fail_msg = "Computer says no. This plot is completely wrong", data_fail_msg = "Make sure you include the data in the ggplot call", aes_fail_msg = "Make sure you define the correct aesthetics (aes)", geom_fail_msg = "Make sure you define the correct plotting function (geom_)")
success_msg("Great!")
```

### 4.1 The `Mean_RT` variable

#### Range, Variance, standard deviation, interquartile range, and  median absolute deviation 
```{r ex="Mean_RT_dispersion", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Mean_RT_dispersion", type="sample-code"}
# The object dataSet is already in your workspace
# Range of Mean_RT

# Variance of Mean_RT

# Standard deviation of Mean_RT

# Interquartile range of Mean_RT

# Median absolute deviation of Mean_RT
```

```{r ex="Mean_RT_dispersion", type="solution"}
# The object dataSet is already in your workspace
# Range of Mean_RT
max(dataSet$Mean_RT, na.rm=TRUE) - min(dataSet$Mean_RT)
# Variance of Mean_RT
var(dataSet$Mean_RT, na.rm=TRUE)
# Standard deviation of Mean_RT
sd(dataSet$Mean_RT, na.rm=TRUE)
# Interquartile range of Mean_RT
IQR(dataSet$Mean_RT, na.rm=TRUE)
# Median absolute deviation of Mean_RT
mad(dataSet$Mean_RT, na.rm=TRUE)
```

```{r ex="Mean_RT_dispersion", type="sct"}
test_output_contains("max(dataSet$Mean_RT, na.rm=TRUE) - min(dataSet$Mean_RT)", incorrect_msg = "Make sure you add na.rm to your max and min statements!")
test_output_contains('var(dataSet$Mean_RT, na.rm=TRUE)', incorrect_msg = "Make sure you add 'na.rm=TRUE' to your code to calculate the variance")
test_output_contains('sd(dataSet$Mean_RT, na.rm=TRUE)', incorrect_msg = "Make sure you add 'na.rm=TRUE' to your code to calculate the standard deviation")
test_output_contains('IQR(dataSet$Mean_RT, na.rm=TRUE)', incorrect_msg = "Make sure you add 'na.rm=TRUE' to your code to calculate the Interquartile range")
test_output_contains('var(dataSet$Mean_RT, na.rm=TRUE)', incorrect_msg = "Make sure you add 'na.rm=TRUE' to your code to calculate the mean absolute deviation")
success_msg("Great job!")
```

#### Box-and-whiskers plot
```{r ex="Mean_RT_boxplot", type="pre-exercise-code"}
library(readr)
dataSet <- read_csv("https://raw.githubusercontent.com/jeroenclaes/statistics_for_linguistics/master/class2_balota_et_al_2007.csv")
```

```{r ex="Mean_RT_boxplot", type="sample-code"}
# The object dataSet is already in your workspace
# Load ggplot2

# Draw a boxplot of Mean_RT

```

```{r ex="Mean_RT_boxplot", type="solution"}
# The object dataSet is already in your workspace
# Load ggplot2
library(ggplot2)
# Draw the boxplot
ggplot(dataSet, aes(x=1, y=Mean_RT)) + geom_boxplot()
```

```{r ex="Mean_RT_boxplot", type="sct"}
test_library_function("ggplot2")
test_ggplot(index=1, all_fail_msg = "Computer says no. This plot is completely wrong", data_fail_msg = "Make sure you include the data in the ggplot call", aes_fail_msg = "Make sure you define the correct aesthetics (aes)", geom_fail_msg = "Make sure you define the correct plotting function (geom_)")
success_msg("Great!")
```

