---
title: 'Topic 3: Exploring qualitative variables'
author: "Jeroen Claes"
output:
  word_document:
    toc: no
  revealjs::revealjs_presentation:
    css: white.css
    self_contained: no
    transition: none
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, message=FALSE, warning=FALSE, error=FALSE, fig.height = 4, dpi = 300)

library(pander)
library(readr)
library(dplyr)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class3_claes_2017.csv")


``` 


## Contents
1. Data for this class
2. Factors
3. Frequencies: counting values
4. Proportions
5. Bar plots
6. References

## 1. Data for this class
- Data from Claes(2017):
    - Corpus investigation into existential agreement variation in Peninsular Spanish
    - Data drawn from *Twitter* and *Corpus Oral y Sonoro del Español Rural (COSER)*
    - Random sample of 500 lines from the dataset
  
```{r data, eval=FALSE}
library(readr)
library(dplyr)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class3_claes_2017.csv")
glimpse(dataSet)
```

## 2. Factors
- When a text/character column can only have a few values (e.g., absent, present) in your analysis, it is best to convert it to a factor
- A factor is a special type of text column:
    - R knows all of its unique values
    - Factors are encoded internally as numbers, and many computations depend on that 
    - Factors have an order, giving you control over the appearance of tables and plots
        - The default order is always alphabetical
    
## 2.1 Conversion from character to factor
- To convert  *character* columns to *factor* you can use `as.factor`
```{r mutateIf1}
dataSet$type <- as.factor(dataSet$type) 
```

## 2.2 Conversion from character to factor
- To convert all *character* values to *factor* we may use the dplyr function `mutate_if`
```{r mutateIf}
library(dplyr)
dataSet <- mutate_if(dataSet, is.character, as.factor)
```

## 2.3 Levels
- The values a factor can assume are called `levels` 
- The first level is called the `reference level` 

```{r levels}
levels(dataSet$type)
```

## 2.4 Re-ordering a factor
- To re-order the levels of a factor, you can use `relevel`

```{r relevel}
levels(dataSet$type)
dataSet$type <- relevel(dataSet$type, ref="singular")
levels(dataSet$type)
```

## 2.4 Re-ordering a factor
- To re-order the levels of a factor based on the values of another column, you can use the function `reorder`
- Useful for organizing the appearance of plots 
```{r reorder1}
levels(dataSet$type)
dataSet$type <- reorder(dataSet$type,dataSet$noun_length)
levels(dataSet$type)
```

## 2.5 Changing factor values
- Often you code your data in a particular way, but during the analysis you realize that your coding is not ideal
- Changing factor values is called `recoding` and it used to be very hard to do
- The function `recode` in the `dplyr` package changed that.
    - Format: `"OldValue"="NewValue"`
```{r recode1}
levels(dataSet$broad.regions)
dataSet$broad.regions <- recode(dataSet$broad.regions, 'North'="Top", "East"="Left", "Center"="Middle", "South"="Bottom") 
levels(dataSet$broad.regions)
```

## 2.5 Changing factor values
- You can also use `recode` to merge factor levels   
```{r recode2}
levels(dataSet$broad.regions)
dataSet$broad.regions <- recode(dataSet$broad.regions, "Left"="Middle") 
levels(dataSet$broad.regions)
```

## 2.6 Excercises
- Go to http://www.jeroenclaes.be/statistics_for_linguistics/labs/class3.html and perform the excercises under *4.Factors*

##  3. Frequencies: counting values
##  3.1 Counting values
- To describe a categorical variable, there's not much you can do except counting how many times each value occurs
- This is what the `table` function does
```{r table}
table(dataSet$type)
```

## 3.2 Counts as data.frames
- Tables are nice for interactive use, but they are less of a tidy data format to program with than data.frames
- Tables can be converted to data.frames with `as.data.frame`
```{r table2}
tab <- as.data.frame(table(dataSet$type))
tab
```

## 3.3 Excercises
- Go to http://www.jeroenclaes.be/statistics_for_linguistics/labs/class3.html and perform the excercises under *5.Frequencies: counting values*

## 4. Proportions
## 4.1 Calculating proportions as decimals
- Proportions are expressed as decimal numbers:
    - `0.25` equals 25%
- Proportions can be calculated from tables with `prop.table`
```{r prop}
prop.table(table(dataSet$type))
```

## 4.2 Proportions as percentages 
- To have the result as numbers from 1 to 100, just multiply by 100
```{r prop2}
prop.table(table(dataSet$type)) * 100
```

## 4.3 Rounding proportions 
- To round the proportions to a number of decimals, we can use `round`
```{r prop21}
tab <- prop.table(table(dataSet$type))  
round(tab, 2)
```

## 4.4 Proportions as data.frames
- You can convert proportions to data.frames, because they are still tables
```{r prop3}
tab <- prop.table(table(dataSet$type)) 
tab <- as.data.frame(tab)
tab 
# In one line:
# tab <- as.data.frame(prop.table(table(dataSet$type))) 

```

## 4.5 Calculating proportions yourself
- Of course, you also store your counts in a data.frame and calculate the proportions yourself
```{r prop4}
tab <- table(dataSet$type)
tab <- as.data.frame(tab)
tab$proportion <- tab$Freq/sum(tab$Freq, na.rm=TRUE)
tab
```

## 4.6 Excercises
- Go to http://www.jeroenclaes.be/statistics_for_linguistics/labs/class3.html and perform the excercises under *6.Proportions*

## 5. Bar plots
## 5.1 Why plots?
- Calculating counts and proportions is great for getting to understand your data, but in papers you will want to use charts:
    - Charts are more intuitive and help people understand the magnitude of the differences between your proportions
- Charts should speak for themselves: 
    - if you can't figure out what the chart is supposed to show by just looking at the plot, you're doing something wrong!

## 5.2 Plotting in `ggplot` (1/6)
- Let's take a moment to review the basic syntax of `ggplot`
- This command sets up the basis of `ggplot`:
    - It defines the data we are working with (`dataSet`)
    - It defines the aesthetics (`aes()`) that are inherited by the other plotting functions that follow behind the `+` signs:
        - Our `x` axis (the horizontal axis) will plot the `type` variable
        - We don't specify a `y` axis, the counts will be calculated by our plotting function
```{r ggplot20, eval=FALSE}
library(ggplot2)
ggplot(dataSet, aes(x=type))
```

## 5.2 Plotting in `ggplot` (2/6)
- After the basic setup, we can add multiple plot layers by adding plotting functions separated by a `+` sign. Here we will consider `geom_bar()`.

```{r ggplot21, eval=FALSE}
library(ggplot2)
ggplot(dataSet, aes(x=type))  + 
  geom_bar()
```

## 5.2 Plotting in `ggplot` (3/6)
- Once we have our plotting layer(s), we can change the appearance of our plot with theme functions:
    - All existing themes are listed on: http://ggplot2.tidyverse.org/reference/ggtheme.html
    

```{r ggplot211, eval=FALSE}
library(ggplot2)
ggplot(dataSet, aes(x=type))  + 
  geom_bar() +
  theme_minimal()
```

## 5.2 Plotting in `ggplot` (3/6)

```{r ggplot211_plot, eval=TRUE}
library(ggplot2)
ggplot(dataSet, aes(x=type))  + 
  geom_bar() +
  theme_minimal()
```


## 5.2 Plotting in `ggplot` (4/6)
- To modify aspects of the theme, we can add a call to `theme`
- Useful theme settings include:
    - `legend.position`:
        - `none`: no legend
        - `bottom`: bottom
    - `legend.title`:
        - `element_blank()`: no (sometimes ugly) legend title
    - `axis.text.x`:
        - `element_text(angle=90)` will rotate your text 90°

```{r ggplot213, eval=FALSE}
library(ggplot2)
ggplot(dataSet, aes(x=type))  + 
  geom_bar() +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle=90), legend.title = element_blank())
```

## 5.2 Plotting in `ggplot` (5/6)
- Finally, we can add axis and plot titles
```{r ggplot2121_code, eval=FALSE}
library(ggplot2)
ggplot(dataSet, aes(x=type))  + 
  geom_bar() +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle=90)) +
  labs(x="Type", y="Frequency", title="haber in Peninsular Spanish")
```
## 5.2 Plotting in `ggplot` (5/6)
```{r ggplot2121, eval=TRUE}
library(ggplot2)
ggplot(dataSet, aes(x=type))  + 
  geom_bar() +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle=90)) +
  labs(x="Type", y="Frequency", title="haber in Peninsular Spanish")
```


## 5.2 Plotting in `ggplot` (6/6)
- Unfortunately, *italicising* parts of (axis) titles is rather complicated 
```{r ggplot212, eval=TRUE, fig.height=3}
library(ggplot2)
titleWithItalics <- expression(paste(italic("Haber "), "in Peninsular Spanish"))
ggplot(dataSet, aes(x=type))  + 
  geom_bar() +
  theme_minimal() + 
  theme(axis.text.x = element_text(angle=90)) +
  labs(x="Type", y="Frequency", title=titleWithItalics)
```

## 5.3 Bar plots
- Barplots are great to visualize counts and proportions
- A simple bar plot can be drawn with the code we already considered
```{r barplot1, eval=FALSE}
library(ggplot2)
ggplot(dataSet, aes(x=type)) + 
  geom_bar()
```

## 5.3.1 Bar plot options (1/3)
- If you set the aesthetics `color` and `fill` to the same variable as your `x` axis, the categories of your variable will be colored differently 
- This will also add a legend
```{r barplotOptions, eval=FALSE}
library(ggplot2)
ggplot(dataSet, aes(x=type, color=type, fill=type)) + 
  geom_bar()
```

## 5.3.1 Bar plot options (1/3)
```{r barplotOptionsPlot, eval=TRUE}
library(ggplot2)
ggplot(dataSet, aes(x=type, color=type, fill=type)) + 
  geom_bar()
```

## 5.3.1 Bar plot options (2/3)
- You can get rid of the legend with `theme()` and the parameter `legend.position` 
```{r barplotOptions21, eval=FALSE}
library(ggplot2)
ggplot(dataSet, aes(x=type, color=type, fill=type)) + 
  geom_bar() + 
  theme(legend.position = "none")
```

## 5.3.1 Bar plot options (2/3)
```{r barplotOptions21Plot, eval=TRUE}
library(ggplot2)
ggplot(dataSet, aes(x=type, color=type, fill=type)) + 
  geom_bar() + 
  theme(legend.position = "none")
```


## 5.3.1 Bar plot options (3/3)
- You can get rid of the legend title with `theme()` and the parameter `legend.title`
```{r barplotOptions2, eval=FALSE}
library(ggplot2)
ggplot(dataSet, aes(x=type, color=type, fill=type)) + 
  geom_bar() + 
  theme(legend.title =  element_blank())
```

## 5.3.1 Bar plot options (3/3)
```{r barplotOptions2Plot, eval=TRUE}
library(ggplot2)
ggplot(dataSet, aes(x=type, color=type, fill=type)) + 
  geom_bar() + 
  theme(legend.title =  element_blank())
```


## 5.3.2 Visualizing proportions with barplots (1/3)
- To visualize proportions with barplots, we have to write some extra code:
    - We set up our ggplot object as we would normally do
    - We add `geom_bar()`
    - We tell `geom_bar()` to calculate its `y` axis by dividing the count `..count..` (this is a special `ggplot` variable) of each value by the sum of the count of the values  `sum(..count..)`

```{r proportions, eval=FALSE}
library(ggplot2)
ggplot(dataSet, aes(x=type))  +
  geom_bar(aes(y=..count../sum(..count..)))
```

## 5.3.2 Visualizing proportions with barplots (1/3)

```{r proportions_plot, eval=TRUE}
library(ggplot2)
ggplot(dataSet, aes(x=type))  +
  geom_bar(aes(y=..count../sum(..count..)))
```


## 5.3.2 Visualizing proportions with barplots (2/3)
- We can tell `ggplot` that the y-axis does not plot decimals, but rather proportions with `scale_y_continuous` and its `label` parameter.
- If we set it to `percent`, then our `y` axis will have nice percentage formatting
- We need the `scales` package for this to work

```{r proportions2a, eval=FALSE}
library(ggplot2)
library(scales)
ggplot(dataSet, aes(x=type))  +
  geom_bar(aes(y=..count../sum(..count..))) +
  scale_y_continuous(labels=percent)
```

## 5.3.2 Visualizing proportions with barplots (2/3)

```{r proportions2aplot, eval=TRUE}
library(ggplot2)
library(scales)
ggplot(dataSet, aes(x=type))  +
  geom_bar(aes(y=..count../sum(..count..))) +
  scale_y_continuous(labels=percent)
```

## 5.3.2 Visualizing proportions with barplots (3/3)
- To make the plot pretty, we can add colors, remove the legend title, and add a better title for the `y` axis

```{r proportions2, eval=FALSE}
library(ggplot2)
library(scales)
ggplot(dataSet, aes(x=type, color=type, fill=type))  +
  geom_bar(aes(y=..count../sum(..count..))) +
  scale_y_continuous(labels=percent) +
  theme(legend.title=element_blank()) + 
  labs(y="Percentages")
```


## 5.3.2 Visualizing proportions with barplots (3/3)

```{r proportions2plot, eval=TRUE}
library(ggplot2)
library(scales)
ggplot(dataSet, aes(x=type, color=type, fill=type))  +
  geom_bar(aes(y=..count../sum(..count..))) +
  scale_y_continuous(labels=percent) +
  theme(legend.title=element_blank()) + 
  labs(y="Percentages")
```

## 5.4 Excercises
- Go to http://www.jeroenclaes.be/statistics_for_linguistics/labs/class3.html and perform the excercises under *7. Bar plots*



## Questions?
- ???

## 6. References
- Claes, J. (2017). Cognitive and geographic constraints on morphosyntactic variation: The variable agreement of presentational haber in Peninsular Spanish. *Belgian Journal of Linguistics* (31), 28-53.