---
title: 'Topic 7: Linear regression: Theory and assumptions'
author: "Jeroen Claes"
output:
  word_document:
    toc: no
  revealjs::revealjs_presentation:
    css: white.css
    self_contained: no
    transition: none
  html_document:
    toc: no
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, message=FALSE, warning=FALSE, error=FALSE, fig.height = 3, dpi = 300)
library(pander)
library(dplyr)
```

## Contents
1. What is linear regression analysis and why is it used?
2. The Least Squares Regression equation
3. Assumptions of regression modeling
4. Validating the assumptions of regression modeling
5. Exercises
6. References

## 1. What is linear regression analysis and why is it used? 
## 1. What is linear regression analysis and why is it used? (1/3)
- We will be working with an extended version of the dataset by Balota et al. (2007) we have seen before (Levshina, 2015):
    - `Word`
    - `Length`: word length
    - `SUBTLWF`:  normalized word frequency 
    - `POS`: part-of-speech
    - `Mean_RT`: Mean reaction time
-  Research question:
    - *To what extent and how do word length, word frequency, and part-of speech condition subjects' reaction times in a lexical decision task?*

```{r data }
library(readr)
library(dplyr)
library(ggplot2)

dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class7_Balota_et_al_2007.csv")
glimpse(dataSet)
```

## 1. What is linear regression analysis and why is it used? (2/3)
- Linear regression is a form of statistical modeling: 
    - We model the values of a continous variable as a function of the values of one or more explanatory variables
        - E.g., Grammatical complexity can be understood as a function of age, speakers vowel quality can be understood as a function of social variables and contextual features,...
    - A linear regression model is a statistical implementation of a series of hypotheses about the way something works
- In its simplest form, linear regression is closely related to correlation analysis:
    - Correlation analysis **describes** the strenghts and significance of the associations between two continuous variables
    - Linear regression attempts to **predict** the values of a continous variable using the values of another continuous variable
  
## 1. What is linear regression analysis and why is it used? (3/3)
- Linear regression has two main purposes:
    - It is used to model a quantity as the additive effect of **several variables**
        - E.g., Reaction times can be modeled as the joint effect of word length, word frequency, age, years of schooling
    - It is used to measure the effect of **one variable** while **controlling** for biases caused by other variables so as to make the effect estimates more meaningful and precise
        - E.g., If we want to investigate the effect of *word length* on reaction times across different subjects, we may want to control for age, word frequency, years of schooling, etc.

## 2. The Least Squares equation
## 2. The Least Squares equation (1/8)
- Linear regression predicts a linear trend line for Y at each point of X
- As you can see, the line does not touch most points.
- The regression equation estimates the **slope** (steepness) that minimizes the squared sum of the differences between the observed values (the dots) and the predicted values (the line):
    - Differences between the predicted and the observed values are called **residuals**
    - The sum of the squared residuals is called the **sum of squares**
    - Linear regression is also called **ordinary least squares** regression
    
```{r scatterplot, eval=TRUE, results='asis', echo=FALSE}
ggplot(dataSet, aes(x=Length, y=Mean_RT)) + 
  geom_point() + 
  geom_smooth(method="lm")
```
 
## 2. The Least Squares equation (2/8)
- Least squares regression is a very powerful technique, but it has some limitations
- In order to truly understand its limitations, it is necessary to have a high-level understanding of what the technique actually does
- Here, we will consider only single-term regression: `Y ~ X`
- In a single-term regression, there are three elements that enter the equation:
    - The `slope coefficient` (the steepness of the line on the plot)
    - The `intercept`: the predicted value of Y in X=0; the Y value where the line crosses/intercepts the Y-axis
    - The predicted value of `Y`
- Let us take a deeper look at each one of these elements
```{r equation, eval=FALSE }
y.predicted <- (slope*X) + intercept
```

## 2. The Least Squares equation (3/8)
- The `slope coefficient` equals the `Pearson product-moment correlation coefficient` of the relationship between X and Y, multiplied by the standard deviation of Y divided by the standard deviation of X
- This brings the correlation coefficient more or less to the same scale as X and Y

```{r equationSlope, eval=TRUE }
slope <- cor(dataSet$Length, dataSet$Mean_RT) * (sd(dataSet$Mean_RT)/sd(dataSet$Length))
slope
```

## 2. The Least Squares equation (4/8)
- The `intercept` equals the average value of Y minus the average value of X multiplied by the slope coefficient
```{r equationIntercept0, eval=TRUE }
intercept <- mean(dataSet$Mean_RT) - (slope*mean(dataSet$Length))
intercept
```

## 2. The Least Squares equation (5/8)
- With these values and our formula we can hand-draw the regression line for X
```{r equationIntercept1, eval=TRUE }
# We generate all values between the min and the max of Length
X<- min(dataSet$Length):max(dataSet$Length)
# We apply our equation
y.predicted <-(slope*X)+intercept
```

## 2. The Least Squares equation (6/8)
- If we store the values in a `data.frame` we can easily plot the observed and the predicted values
```{r equationIntercept, eval=TRUE }
values <- data.frame(x=X, y=y.predicted)

ggplot(dataSet, aes(x=Length, y=Mean_RT)) + 
  geom_point() + 
  geom_line(data=values, aes(x=x, y=y), color="red")
```

## 2. The Least Squares equation (7/8)
- With more than one predictor variable, the equation remains identical, but calculating the intercept becomes more complicated: 
    - In `multiple regression` or `multivariate regression` the intercept equals the value of the dependent variable when all predictor variables are zero
    - The predicted value is calculated as the sum of the intercept and the slope estimate for each variable
    - The geometric form of the equation is no longer a line, but a *hyperplane*, a k-dimensional shape in k-dimensional space
- In most cases you will want to construct a multivariate model to parcel out the effects of each individual variable

```{r equationMultiplePredictor2s, eval=FALSE }
y.predicted <- intercept + (slope1*dataSet$Length) + (slope2*dataSet$SUBTLWF) +  ...
```

## 2. The Least Squares equation (8/8)
- The `lm` ("linear model") function in R will take care of calculating the intercept, slope and predicted values for us
- It uses `formula notation`:
    - The leftmost string is the dependend variable (`Mean_RT`)
    - The `tilde` marks that the right-hand variable names are independent variables
    - The `+` signs refer to the additive effect of the variables in the least squares equation
- Don't forget to specify the data.frame that holds your data!
- If your model includes all the columns in your data.frame, you can use short-hand notation:
    - `Mean_RT ~ .`

```{r equationMultiplePredictors, eval=TRUE }
mod <- lm(Mean_RT ~ Length +   SUBTLWF + POS, data = dataSet)
```

## 3. Assumptions of linear regression modeling
## 3. Assumptions of linear regression modeling (1/4)
- The least-squares regression equation immediately exposes some of its assumptions and limitations:
    - The slope is calculated using the Pearson product-moment correlation coefficient: 
        - The technique will inherit the assumptions of the Pearson product-moment correlation coefficient (see Topic 5):
            - Y is interval- or ratio-scaled (there are ways of dealing with ordinal or categorical data in X)
            - The relationship between X and Y is **monotonic** 
            - The relationship between X and Y is **linear**
            - There are **no outliers** in X and Y
            
## 3. Assumptions of linear regression modeling (2/4)
- The least-squares regression equation immediately exposes some of its assumptions and limitations:
    - The slope is calculated using the standard deviations of X and Y:
        - X and Y should follow the `normal distribution`, otherwise their standard deviations will be biased and the resulting slope estimate will be biased as well
    - The intercept is calculated using the `means` of X and Y:
        - X and Y will have to be normally distributed, or the means will not be good indicators of their central tendencies, resulting in poor intercept estimates
        
## 3. Assumptions of linear regression modeling (3/4)
- The least-squares regression equation immediately exposes some of its assumptions and limitations:
    - The predicted value of `Y` is calculated by adding up `(slope*Predictor1) + (slope*Predictor2) ...`:
        - The effects of the independent variables should be additive, this means that the independent variables cannot be strongly correlated (there is no `multicollinearity`)

## 3. Assumptions of linear regression modeling (4/4)
- Levshina (2015: 155) lists the following additional assumptions:
    - The observations are independent from one another (there is no `autocorrelation`)
    - The variability of the residuals does not increase or decrease with the explanatory variables or the response (i.e., there is no `heteroskedasticity`)
    - The residuals are not autocorrelated
    - The residuals follow the normal distrubtion. This is less important if the sample size is large 

## 4. Validating the assumptions of regression modeling
## 4. Validating the assumptions of regression modeling
- Before you think of fitting a linear regression model, you have to:
    - Investigate the distributions of all of your variables:
        - What are their shapes: are they normally distributed?
        - What are their central tendencies?
        - What is their dispersion?
        - Are there any outliers?
        
## 4. Validating the assumptions of regression modeling
- Before you think of fitting a linear regression model, you have to:
    - Examine the relationship of the independent variables to your dependent variable:
        - Are they linear?
        - Are they homeoskedastic? 
    - Examine the relationships between each of your individual predictors:
        - Are they correlated? To which extent? 
- After checking all of this, you can specify a first regression model to validate the regression assumptions further

## 4.1 The observations are independent from one another 
- Each value of the response variable should be independent from the values of other observations:
    - E.g., the reaction times for *marveled* should not depend on the reaction times for another word
    - For study designs where you gather multiple observations for the same subject/word, a `mixed-effects regression model` (a model that takes this fact into account) is more appropriate than the simple linear model

## 4.2 There are no outliers/overly influential observations
- Outliers may distort the regression analysis and should be removed:
    - A boxplot of each of your variables will help you identify outliers
- A `density plot` or a `QQ-plot` will help you estimate the shape of the distribution of each variable
    - Depending on the shape of the distribution of your variable, you may want to:
        - For normally distributed variables: 
            - Remove outliers based on z-scores 
        - For data that does not follow the normal distribution: 
            - Remove outliers based on MAD-scores or the interquartile range 
            
## 4.3 There are no outliers/overly influential observations
- Remove outliers based on z-scores:
    - Here our dependent `Mean_RT` variable and our `Length` variable are mostly normally distributed, except for a couple of atypical observations
    - Values that provide an absolute z-score of more than 2 are outliers (2 standard deviations higher/lower than the mean)
```{r outliersZscores, eval=TRUE }
dataSet<-dataSet[abs(scale(dataSet$Mean_RT)) <= 2, ]
dataSet<-dataSet[abs(scale(dataSet$Length)) <= 2, ]
```        

## 4.4 There are no outliers/overly influential observations
- Remove outliers based on MAD-scores/interquartile range:
    - The `SUBTLWF` variable has the typical power-law distribution of word frequencies
    - Here we have two possibilities:
        - We exclude all variables that have a value smaller/higher than the first/third quartile minus/plus `1.5 times the interquartile range`. Note that we have to use `which` to find those values, because we need to specify an `AND` statement (written as `&`)
        - We exclude all variables that have an absolute value greater than `3.5` when the median is substracted from the value and divided by the median absolute deviation 
        
```{r outliersMad, eval=TRUE }
dataSet <- dataSet[abs((dataSet$SUBTLWF-median(dataSet$SUBTLWF))/mad(dataSet$SUBTLWF)) <= 3.5, ]
# dataSet<- dataSet[which(dataSet$SUBTLWF >= quantile(dataSet$SUBTLWF, 0.25) - (1.5*IQR(dataSet$SUBTLWF)) & 
# dataSet$SUBTLWF <= quantile(dataSet$SUBTLWF, 0.75) + (1.5*IQR(dataSet$SUBTLWF)) ),]
# We refit our model to the changed data
mod <- lm(Mean_RT ~ Length +   SUBTLWF + POS, data = dataSet)
```        

## 4.4 There are no outliers/overly influential observations
- After these steps, we may define an intial regression model
- The `influencePlot` function from the `car` package identifies the row indices of too influential observations

## 4.4 There are no outliers/overly influential observations
- The plot combines three values:
    - Studentized residuals (Y-axis). These values represent the difference between observed and predicted values. Values more extreme than `2` should be investigated
    - Hat-values or leverage (X-axis). How much influence an observation may have on the predicted values  
    - Cook's distance (size of the bubbles). How much influence removing an observation may have on the regression coefficients

```{r outliersInfluencePlot, eval=TRUE }
library(car)
influencePlot(mod)
```

## 4.4 There are no outliers/overly influential observations
- The function returns a data.frame with the statistics for the most suspicious observations
- The `row.names` of the data.frame are the indices of the observations that should be removed
- It is usually a good idea to remove these

```{r outliers2, eval=TRUE }
library(car)
indices <-row.names( influencePlot(mod))
indices <-as.numeric(indices)
dataSet <- dataSet[-indices, ]
# We refit our model to the changed data
mod <- lm(Mean_RT ~ Length +   SUBTLWF + POS, data = dataSet)
```

## 4.5 The relationship between the dependent and independent variables is linear
- The first thing you should do is plotting the relationship of your numeric predictors to the dependent variable to fix the easy to spot non-linear trends
- For regression models the `crPlot` function in the `car` package plots the relationship between X and Y
- There are two lines on the plot: 
    - The dashed red line plots the expected linear relationship
    - The green line plots the observed relationship
    
```{r linearity, eval=TRUE }
library(car)
crPlot(mod, var = "Length")
```

## 4.5 The relationship between the dependent and independent variables is linear
- If the two lines diverge greatly, the relationship is not linear
- The relationships are not perfectly linear, but they come pretty close
- Non-linear relationships may be made linear with transformations of the predictor variable
- In other cases, it is more appropriate to model them with `polynomial regression terms`, linear approximations of the curvy relationship (see Topic 12). This makes the model less easily interpretable

```{r linearity2, eval=TRUE }
library(car)
crPlot(mod, var = "SUBTLWF")
```

## 4.6 There is no heteroskedasticity
- The previous two plots reveal that there is a bigger issue than some non-linearity: heteroskedasticity. Both for the `Length` and the  `SUBTLWF` variable the plots display a funnel-like pattern
- This means that the variability of the residuals covaries (i.e, increases/decreases) with the explanatory variables or the response

## 4.6 There is no heteroskedasticity

```{r heteroskedasticity, eval=TRUE }
plot(mod, which = 1)
```

## 4.7 There is no heteroskedasticity
- The `car` package includes the `ncvTest` function
- If p < 0.05, the fitted (predicted) values have non-constant variance with the residuals 
- This test establishes whether or not heteroskedasticity is an issue for the model
- Here it is clear that our model suffers from heteroskedasticity 
```{r heteroskedasticity2, eval=TRUE }
library(car)
ncvTest(mod)
```

## 4.7 There is no heteroskedasticity
- We should also check whether the residuals have constant variance depending on the values of quantitative explanatory variables
- We perform the same  `ncvTest`  test, but we add the predictor variable 
- This allows us to check if the problem is due to specific variables. Here we find that the `SUBTLWF` variable is the main culprit

```{r heteroskedasticity3, eval=TRUE }
library(car)
ncvTest(mod, ~ Length)
ncvTest(mod, ~ SUBTLWF)
ncvTest(mod, ~POS)
```

## 4.7 There is no heteroskedasticity
- Heteroskedasticity can be fixed with a power transformation of the **dependent variable**
- We can try to find the right transformation by trial and error, or we can use a `boxCox` plot (from the `car` package) 
- The X-value with the highest Y-value corresponds to the power of the optimal transformation for our dependent variable
- If the Y-value is zero, the optimal transformation is the natural logarithm
```{r heteroskedasticity4, eval=TRUE }
library(car)
boxCox(mod)
```

## 4.7 There is no heteroskedasticity
- We can find the exact x-value that corresponds with the highest y-value as follows:
    - We store the data for the plot in an object
    - We set `plotit` to `FALSE` so we don't get a plot output
    - We subset the x-vector by calling `which.max` (which vector item has the highest value) on the y-values
```{r heteroskedasticity5, eval=TRUE }
library(car)
testValues <- boxCox(mod, plotit=FALSE)
power <- testValues$x[which.max(testValues$y)]
power
```

## 4.7 There is no heteroskedasticity
- All we have to do now to fix the problem is raising `Mean_RT` to the power of ``r power`` and compute the model again
- When we perform the `ncvTest` again, the  p-value is higher than 0.05, indicating that heteroskedasticity is no longer an issue

```{r heteroskedasticity6, eval=TRUE }
library(car)
dataSet$Mean_RT <- dataSet$Mean_RT^power
mod <- lm(Mean_RT ~ Length +  SUBTLWF, data = dataSet)
ncvTest(mod)
```

## 4.8 The explanatory variables are not correlated (there is no `multicollinearity`)
- For numerical variables:
    - Perform pairwise correlation tests (see Topic 5)
    - You will want to see p-values higher than 0.05 and/or weak to moderate correlations
- For categorical variables:
    - Perform pariwise Chi-squared tests of independence and calculate the effect sizes (see Topic 6)
    - You will want to see p-values higher than 0.05 and/or small to moderate effect sizes
- You can also assess multicollinearity directly from a fitted regression model:
    - Calculate the Variance Inflation Factors (VIF)
    - `VIF <= 5` indicates that there is no severe multicollinearity 
    - `vif` in the `car` package
    
```{r multicol, eval=TRUE }
library(car)
vif(mod)
```

## 4.8 The explanatory variables are not correlated (there is no `multicollinearity`)
- If some of your variables are collinear you should drop one of them
- If it is of theoretical interest to investigate both variables, you can run parallel regression models, which each include one of the variables
- If interpretability is not of primary concern (say, when the collinear variables only control for biases), techniques such as `principal components` (for quantitative variables) or `factor analysis` (for factors) can combine the two collinear variables into one or more non-correlated variables 

## 4.9 The residuals are not autocorrelated
- Just like the observations cannot be autocorrelated, the residuals cannot be autocorrelated either
- To detect autocorrelation, we can perform a `durbinWatsonTest`
- `durbinWatsonTest` in the `car` package 
- If p > 0.05 autocorrelation is not an issue

```{r autocor, eval=TRUE }
library(car)
durbinWatsonTest(mod)
```

## 4.9 The residuals are not autocorrelated
- If the residuals are autocorrelated this may mean that:
    - We did not include a variable that ought to be included
        - **Solution**: Think, plot, and add the predictor
    - We misspecified the shape of some critical relationship. This happens when we fit a straight line for a predictor that has a non-linear relationship to the dependent variable. 
        - **Solution**: Add a non-linear (polynomial) term (see Topic 12)
    - We have made some consistent error while measuring the dependent variable
        - **Solution**: Find the observations with errors and exclude them
- Autocorrelation will not affect the sizes and the numeric stability of the regression coefficients, but it will affect their standard errors, t-values, and p-values
- Autocorrelation will also affect the predictive accuracy of your model

## 4.10 The residuals follow the normal distribution
- If everything is in order, the residuals (the "prediction errors" of our model) should be completely random, resulting in normally distributed values
- If the residuals are not normally distributed, this could indicate that our model misses important information 
    - E.g., We may have failed to include an important variable 
- To inspect the residuals you can simply draw a `qqplot` with a `qqline` of the residuals (`resid` function). 
- Alternatively, you can draw the plot with `plot(mod, which = 2)`, which adds index labels for problematic observations

```{r residuals, eval=TRUE }
qqnorm(resid(mod)) 
qqline(resid(mod))
```

## 5. Exercises
- Please go to http://www.jeroenclaes.be/statistics_for_linguistics/class7.html and perform the exercises

## 6. References
- Balota, D.A., Yap, M.J., & Cortese, M.J., et al. (2007). The English Lexicon Project. *Behavior Research Methods*, 39(3), 445–459. DOI: 10.3758/BF03193014. Data taken from Levshina (2015).
- Levshina, N. (2015). *How to do Linguistics with R: Data exploration and statistical analysis*. Amsterdam/Philadelphia, PA: John Benjamins.
- Urdan, T. (2010). *Statistics in Plain English*. New York NY/London: Routledge.
