---
title: 'Topic 6: Relationships between qualitative variables'
author: "Jeroen Claes"
output:
  word_document:
    toc: no
  revealjs::revealjs_presentation:
    css: white.css
    self_contained: no
    transition: none
  html_document:
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=TRUE, message=FALSE, warning=FALSE, error=FALSE, fig.height = 3, dpi = 300)
library(pander)
library(dplyr)
library(ggplot2)
```

## Contents
1. Cross-tabulation: the basic tool to explore relations between qualitative variables
2. Chi-square test of independence
3. Fisher exact test
4. Effect sizes for relations between qualitative variables
5. Plots to explore the associations between qualitative variables 
6. Reporting on associations between qualitative variables
7. Exercises
8. Questions
9. References

## 1. Cross-tabulation: the basic tool to explore relations between qualitative variables
## 1. Cross-tabulation: the basic tool to explore relations between qualitative variables
## 1.1 Data
- We will be working with a dataset by Claes (2017)
    - Corpus investigation into existential agreement variation in Peninsular Spanish
    - Data drawn from *Twitter* and *Corpus Oral y Sonoro del Español Rural (COSER)*
    - Random sample of 500 lines from the dataset
  
```{r data, eval=TRUE}
library(readr)
library(dplyr)
dataSet <- read_csv("http://www.jeroenclaes.be/statistics_for_linguistics/datasets/class3_claes_2017.csv")
head(dataSet)
```

## 1.2 Crosstabulation (1/2)
- To explore the relationship between two qualitative variables, we have to count how many times each combination of the two variables occur
- This is called `crosstabulation`
- The output table is called a `contingency table`
- This can be done with the `table` function
- So, to count the number of times the `singular` and the `plural` existential occur in each of the `negation` groups, we use:
```{r table1, eval=TRUE}
# 'x' axis first (groups/independent variable), then 'y' axis (dependent variable)
table(dataSet$negation, dataSet$type)
```

## 1.2 Crosstabulation (2/2)
- Absolute counts are nice, but we're usually interested in the proportions of one option vs another
- This is what `prop.table` does. It takes two arguments:
    - A table
    - A `margin`:
        - `1` calculate the proportions by row (each **row** sums 1)
        - `2` calculate the proportions by column (each **column** sums 1)

```{r table2, eval=TRUE}
# 'x' axis first (groups/independent variable), then 'y' axis (dependent variable)
tab<-table(dataSet$negation, dataSet$type)
# proportions by row
prop.table(tab, 1)
# proportions by column
prop.table(tab, 2)
```

## 2. Chi-square test of independence
## 2. Chi-square test of independence 
- The table shows that there is a sizeable difference between the frequencies of the singular and the plural existential in clauses with and without negation
- But is this difference reliably large? To test this, we need to perform a `Chi-squared test of independence`
- The `null hypothesis` of this test is that there is no association between the variables:
    - *There is no influence of the presence of negation on the occurrence rate of the singular and the plural existential*
- The test compares the `observed frequency` to the frequency you would expect by chance (`the expected frequency`) to see how likely the results are

## 2. Chi-square test of independence 
- To get the expected frequency for a table cell, we need to add the `marginal frequencies` to the table
- Marginal frequencies are the row and column totals
- We can get the marginal frequencies with `rowSums` and `colSums`
```{r table3, eval=TRUE}
tab <- table(dataSet$negation, dataSet$type)
colSum<- colSums(tab)
rowSum<-rowSums(tab)
```

## 2. Chi-square test of independence 
- For the cell in column 1, row 1, the expected frequency can be found with the following formula:
```{r table4, eval=TRUE}
sampleSize <- sum(tab)
(colSum[1]/sampleSize) * (rowSum[1]/sampleSize) * sampleSize
```

## 2. Chi-square test of independence 
- Basically, the chi-square test functions as follows:
    - Calculate the expected frequency for each cell
    - Substract the expected frequency from the observed frequency. The resulting number is the `residual`
    - Divide this number by the square root of the expected frequency. The resulting number is the `Pearson residual`
    - Square all Pearson residuals and sum them. The resulting number is the `Chi-squared statistic`. 
    - Look up the chi-squared value in a chi-squared probability table. The degrees of freedom are the number of rows in the table, minus one.
- The R function `chisq.test` does all of that for us

## 2.1 Assumptions of the Chi-square test of independence
- The sample is randomly selected from the population of interest and the observations are independent. 
- Every observation can be classifed into exactly one category according to the criterion represented by each variable (Levshina, 2015: 212)
- The expected frequency for all cells is greater than 5
- The numbers in the table are counts, not proportions (**NEVER** run association tests on percentages!)

## 2.2 Performing a Chi-square test of independence in R (1/4)
- The function `chisq.test` takes two factor or character columns as its arguments
- It also accepts a matrix or a table of numeric values
- You don't have to specify the directionality of the hypothesis, because it is always assumed to be two-tailed
```{r chisq.1, eval=TRUE}
chisq.test(dataSet$negation, dataSet$type)
```
 
## 2.2 Performing a Chi-square test of independence in R (2/4)
- If you run the test on proportions, you end up with a p-value of 1
```{r chisq.3, eval=TRUE}
chisq.test(prop.table(table(dataSet$negation, dataSet$type),2))
```

## 2.2 Performing a Chi-square test of independence in R (3/4)
- If you want to inspect the `expected frequencies`, you can pull them out of the test (or calculate them by hand with the formula above)
- Remember, if your expected frequency in one of the cells is lower than 5, the Chisq-test is not appropriate.
- R will fire the following warning when this is the case:
    - `Chi-squared approximation may be incorrect`
```{r chisq.21, eval=TRUE}
chisq.test(dataSet$negation, dataSet$type)$expected
```

## 2.2 Performing a Chi-square test of independence in R (4/4)
- The `chisq.test` also provides access to the residuals, the difference between the predicted and the expected values
- To compare the residuals in a meaningful way, they are standardized by dividing the residuals by the square root of their variance
- If we know the values of the standardized residuals, we may gain insight into which cells contributed significantly to the p-value
- `Standardized residuals` greater than `1.96`/smaller than `-1.96` indicate that a cell contributed significantly to the chi-squared value at the level of 0.05
- If only one "odd" table cell contributes nearly all of the chi-squared (without you expecting it to do so on theoretical grounds), there may a data imbalance issue you will have to investigate further. This is more likely to occur in larger-dimensionality tables
```{r chisq.2, eval=TRUE}
chisq.test(dataSet$negation, dataSet$type)$stdres
```    

## 3 Fisher-Yates exact test
## 3 Fisher-Yates exact test
- When one of your expected frequencies is lower than 5, it is better to perform a Fisher-Yates exact test
```{r fisherYates, eval=TRUE}
fisher.test(dataSet$negation, dataSet$type)
```

## 4. Effect sizes for relations between qualitative variables
## 4. Effect sizes for relations between qualitative variables (1/2)
- The p-values only tell you if the differences between the groups created by the X-variable is reliably large
- They tell you nothing about the sizes of the differences
- In the `vcd` package, the `assocstats` function computes several effect size statistics you can use to measure the strength of the differences
- They are computed on a `table`
```{r assocstats, eval=TRUE}
library(vcd)
tab<-table(dataSet$negation, dataSet$type)
assocstats(tab)
```

## 4. Effect sizes for relations between qualitative variables (2/2)
- For 2x2 tables, the measures are all identical (but not for e.g., 4x2 tables!)
    - Cramér’s *V* 
    - *ϕ* coefficient (phi) 
    - Contingency coefficient (used less often)
- For 2x2 tables, Cramér’s *V* and *ϕ*  range from 0 to 1
- For larger tables,  *ϕ* surpasses 1, so the function only returns Cramér’s *V* in that case (*ϕ* is `NA`)
- Rule of thumb for Cramer's *V* (and  *ϕ* in 2x2 tables):
    - 0: no association
    - 0-0.3: small effect size
    - 0.3-0.5: moderate effect
    - 0.5-0.99: large effect
    - 1: perfect association
    
```{r assocstats2, eval=TRUE}
library(vcd)
tab<-table(dataSet$province, dataSet$type)
assocstats(tab)
```

## 5. Plots to explore the associations between qualitative variables 
## 5. Plots to explore the associations between qualitative variables (1/5)
- Barplots are the method of choice to plot associations between qualitative variables
- They come in three flavors:
    - Side-by side: (`position="dodge"`)
    - Stacked: (`position="stack"`)
    - Percentages: (`position="fill"`)
- Observe that we do not specify a `y`axis mapping here
```{r dodge, eval=TRUE}
ggplot(dataSet, aes(x=negation, fill=type, color=type)) + geom_bar(position = "dodge")
``` 

## 5. Plots to explore the associations between qualitative variables (2/5)
- Barplots are the method of choice to plot associations between qualitative variables
- They come in three varieties:
    - Side-by side: (`position="dodge"`)
    - Stacked: (`position="stack"`)
    - Percentages: (`position="fill"`)
- Observe that we do not specify a `y`axis mapping here
```{r stack, eval=TRUE}
ggplot(dataSet, aes(x=negation, fill=type, color=type)) + geom_bar(position = "stack")
``` 

## 5. Plots to explore the associations between qualitative variables (3/5)
- Barplots are the method of choice to plot associations between qualitative variables
- They come in three varieties:
    - Side-by side: (`position="dodge"`)
    - Stacked: (`position="stack"`)
    - Percentages: (`position="fill"`)
- Observe that we do not specify a `y`axis mapping here
- We add `scale_y_continous` to have nice percent-formatted y-axis labels (requires scales package)
```{r fill, eval=TRUE}
library(scales)
ggplot(dataSet, aes(x=negation, fill=type, color=type)) + 
  geom_bar(position = "fill") + 
  scale_y_continuous(labels = percent)
``` 

## 5. Plots to explore the associations between qualitative variables (4/5)
- We can add data labels to our `dodged` plots with a text layer
- For `stacked` and `filled` plots, this is very complicated

```{r stack0, eval=TRUE}
ggplot(dataSet, aes(x=negation, fill=type, color=type, label=..count..)) + 
  geom_bar(position = "dodge") + 
  geom_text(stat='count', position=position_dodge(width = 1), vjust=-1, color="black")
``` 

## 5. Plots to explore the associations between qualitative variables (5/5)
- `stacked` plots are great to show your readers how much tokens you have for each 'condition' (e.g., for negation, for different locations, ...)
- `dodged` plots are great to compare raw frequencies
- `filled` plots are preferred when you want to make the point that the relative frequency of your dependent variable varies by condition

## 6. Reporting associations between qualitative variables
## 6. Reporting associations between qualitative variables
- Contingency table
- Test type (Chi-squared, fisher's exact)
- Test statistic (Chi-squared)
- Degrees of freedom
- Effect sizes
- Plots

## 7. Exercises
- Please go to: http://www.jeroenclaes.be/statistics_for_linguistics/class6.html and perform the exercises. 

## 8. Questions?
???

## 9. References
- Claes, J. (2017). Cognitive and geographic constraints on morphosyntactic variation: The variable agreement of presentational haber in Peninsular Spanish. *Belgian Journal of Linguistics* (31), 28-53.
- Levshina, N. (2015). *How to do linguistics with R: Data exploration and statistical analysis*. Amsterdam/Philadelphia, PA: John Benjamins.
- Urdan, T. (2010). *Statistics in Plain English*. New York NY/London: Routledge.